### 2019-08-04 Thesis/pmonads2.tex
Put everything into one document, added title page, did spell check etc.

\documentclass{report}[11pt]
\let\FEWFONTS=1

\input{../common/preamble}
\usepackage{lua-visual-debug}

\begin{document}

@@

\newcommand{\IAX}{{IA${}_{X}$}\xspace}
\newcommand{\IAXX}{{IA${}_{\X}$}\xspace}
\section{The language \IAXX}

@@

Let $\G$ be a model of Idealized Algol as described above, and suppose that there is an oplax monoidal functor $\Set \to \G$ that is used to interpret datatypes.  
We will use an underline to indicate thies functor; so, for example, the object of $\G$ that is used to denote the natural number type is written $\ul\bN$.

By our description of $\G/\oppcat \X$ as a lax colimit in $\Cat$ (i.e., Corollary \ref{CorTheConstructionUniversalProperty}), we have a natual functor $J\from \G\to\G/\oppcat \X$ and a natural transformation $\phi_{p,a}\from J(jp\to a) \to Ja$.  
Our denotational semantics of \IAXX is then given in the category $\G/\oppcat X$ as follows.  
The denotation of any type $T$ of Idealized Algol is given by $J(\deno{T}_\G)$, where $\deno{T}_\G$ is the original denotation in $\G$.
The denotation of any sequent $\Gamma\ts M$ is given by $\deno{\Gamma\ts M} = J(\deno{\Gamma\ts M}_\G)$, where $\deno{-}_\G$ is the original denotation in $\G$.
The denotation of $\choose_p$ is given by the morphism $\omega_p \from 1 \to j(p)$ given by the composite
\[
  I \xrightarrow{\Lambda(\id_{Jjp})} (Jjp \to Jjp)  \to J(jp \to jp) \xrightarrow{\phi_{p,jp}} Jjp\,.
  \]

@@

Recall that, in the definition of the category $\G/\oppcat\X$, the \Mellies morphisms are left unchanged by precomposing with a morphism in the image of the functor $j$; therefore, if $M\from T$ is a closed term of \IAXX referring to $\choose_{p_1},\cdots,\choose_{p_n}$, then we may write the denotation of $M$ as the composite
\[
  j(N) \xrightarrow{jf} j(p_1 \tensor \cdots \tensor p_n) \xrightarrow{m^j} j(p_1) \times \cdots \times j(p_n) \xrightarrow{\deno{x_1,\cdots,x_n\ts M[x_i/\choose_{p_i}]}_\G} \deno T\,;
  \]
i.e., as
\[
  j(N) \xrightarrow{\langle \pi_1,\cdots,\pi_n\rangle} j(p_1) \times \cdots \times j(p_n) \xrightarrow{\deno{x_1,\cdots,x_n\ts M[x_i/\choose_{p_i}]}_\G} \deno T\,.
  \]

@@

\begin{lemma}
  Let
  \[
    \inferrule{\Gamma,s^{(0)}\ts M_1\converges_{u_1} c_1,s^{(1)} \\ \cdots \\ \Gamma,s^{(n-1)}\ts M_n\converges_{u_n} c_n,s^{(n)}}
    {\Gamma,s^{(0)} \ts M \converges_{u_1\cat \cdots \cat u_n} c,s^{(n)}}
    \]
  be an inference of IA${}_{j(N)}$, where every instance of $\ask_{j(N)}$ occurs as part of some term $\neww (\lambda v.v\gets \ask_{j(N)};\Pi_i \oc v)$, and suppose that $M\ne \neww(\lambda v.v\gets\ask_{j(N)};\Pi_j\oc v)$ for any $j$.  
  Suppose we have sequences $U_1,\cdots,U_n$ such that $u_k$ covers $U_k$ for $k=1,\cdots,n$.
  Then
  \[
    \inferrule{\Gamma,s^{(0)}\ts M_1[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_1} c_1,s^{(1)} \\ \cdots \\ \Gamma,s^{(n-1)}\ts M_n[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_n} c_n,s^{(n)}}
    {\Gamma,s^{(0)} \ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_1\cat \cdots \cat U_n} c,s^{(n)}}
    \]
  is a valid inference of \IAXX.
  \label{LemThirdSubstitution}
\end{lemma}
\begin{proof}
  As in Lemma \ref{LemSecondSubstitution}, we can prove this by looking at cases.  
  For example, consider the sequencing rule
  \[
    \inferrule{\Gamma,s\ts M\converges_u \skipp,s' \\ \Gamma,s' \ts N \converges_v c,s''}{\Gamma,s\ts M;N \converges c,s''}\,.
    \]
  We have
  \begin{IEEEeqnarray*}{Cl}
    &(M;N)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \\
    = &M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)];\\&N[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)]\,,
  \end{IEEEeqnarray*}
  and so we certainly get a rule
  \[
    \inferrule*{\Gamma,s\ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)]\converges_U \skipp,s' \\ \Gamma,s' \ts N[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_V c,s''}{\Gamma,s\ts (M;N)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges c,s''}\,.
    \]
  The only case where we need to be careful is for the $\neww$ rule:
  \[
    \inferrule*{\Gamma,x,(s\vert x\mapsto 0)\ts M\converges_u c,(s'\vert x\mapsto n)}
    {\Gamma,s\ts \neww \lambda x.M \converges_u c,s'}\,.
    \]
  If $\neww\lambda x.M \ne \neww(\lambda v.v\gets \ask_{j(N)};\Pi_j \oc v)$, then we have
  \begin{IEEEeqnarray*}{Cl}
    &(\neww \lambda x.M)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \\
    = & \neww \lambda x. (M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)])\,.
  \end{IEEEeqnarray*}
  Then we can apply the rule
  \[
    \inferrule*{\Gamma,x,(s\vert x\mapsto 0)\ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_U c,(s'\vert x\mapsto n)}
    {\Gamma,s\ts (\neww\lambda x.M)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges U c,s'}.\qedhere
    \]
\end{proof}

@@

\begin{lemma}
  Let $\Gamma,y_1\from j(p_1),\cdots,y_n\from j(p_n) \ts M \from T$ be a term-in-context of ordinary Idealized Algol, where $\Gamma$ is a $\Var$-context.  
  Let $U$ be a sequence and let $N,\pi_i,\Pi_i$ be as above.
  Suppose that there exists some sequence $u\in j(N)^*$ such that $u$ covers $U$ and such that
  \[
    \Gamma,s\ts M[\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i \oc v)/y_i]\converges_u c,s'\,.
    \]
  Then
  \[
    \Gamma,s\ts M[\choose_{p_i}/y_i] \converges_U c,s'\,.
    \]
  \label{LemAdequacyIaxx}
\end{lemma}
\begin{proof}
  Induction on the derivation.
  Suppose that $M$ is not one of the $y_i$; then $M[\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)/y_i]$ is not equal to $\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)$.
  Moreover, every instance of $\ask_{j(N)}$ in $M$ occurs as part of an expression of the form $\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)$, and so we win by Lemma \ref{LemThirdSubstitution} and the inductive hypothesis.

@@

Given a closed term $M\from\com$ of IA${}_\bB$, we define
\[
  \bP(M\converges) = \sum_{u\in\bB^*\colon M\converges_u\skipp} 2^{-|u|}\,,
  \]
since $2^{-|u|}$ is the probability of a particular sequence $u$ of $\true$ and $\false$ values occurring.
Here, the infinite sum means the supremum over all sums of finite subsets.
Similarly, given a Kleisli morphism $\sigma\from 1 \to \bC$ -- i.e., a morphism $\sigma\from 1 \to (\bB \to \bC)$ in $\G$, we can define
\[
  \bP(\sigma\downarrow) = \sum_{u\in\Acc(\sigma)} 2^{-|u|}\,.
  \]
Our Computational Adequacy result for IAX (Propositions \ref{PropKleisliSoundness} and \ref{PropKleisliAdequacy}) then gives us a Computational adequacy result for this model.

@@

This model is much easier to realize.  
What do we gain, then, from our more complicated model?

The most obvious answer is that our original model allowed us to work with arbitrary probabilities, rather than using a fixed coin with probability $\frac12$.  
This is not such a great advantage as it might seem, since if $p\in[0,1]$ is any real number whose binary expansion is computable as a function $\bN \to \bB$, then we can simulate $\choose_p$ within the probabilistic version of IA${}_\bB$\footnote{Idea: build up a sequence of intervals $I_n$ of width $2^n$ by repeatedly flipping the coin and using the result to choose either the lower or the upper half of $I_{n-1}$.  
If at any point the upper bound of $I_n$ is less than or equal to $p$ (which can be checked by computing the first $n$ terms of the binary expansion of $p$, then return $\true$.  
If at any point the lower bound of $I_n$ is greater than or equal to $p$, return $\false$.}.

@@

\begin{proposition}
  Let $\sigma\from A \to B$, $\tau\from B \to C$ be morphisms in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$.  
  Let $s$ be a legal play in $A \to C$.
  Then
  \[
    \bP(s\in\sigma;\tau) = \sum_{\s\in\wit_B(s)}\bP(\s\vert_{A,B}\in\sigma)\bP(\s\vert_{B,C}\in\tau)\,.
    \]
  \label{PropCompositionProb}
\end{proposition}
\begin{proof}
  Suppose that $\sigma$ and $\tau$ are given by (equivalence classes of) pairs
  \begin{mathpar}
    (V\from\Omega\to X, \tilde \sigma \from A \to (\ul X \to B))
    \and
    (W\from\Omega\to Y, \tilde \tau \from B \to (\ul Y \to C))\,,
  \end{mathpar}
  where $V$ and $W$ are random variables and $\tilde \sigma,\tilde\tau$ are strategies in $\G$.
  Then the composition $\sigma;\tau$ in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$ is given by $V\tensor W\from\Omega \to X\times Y$, together with the \Mellies composition
  \[
    A \xrightarrow{\tilde\sigma}
    (\ul{X} \to B) \xrightarrow{\ul{X} \to \tilde\tau}
    (\ul X \to (\ul Y \to C)) \to
    ((\ul X \times \ul Y) \to C) \xrightarrow{\mu\to C}
    (\ul{X\times Y} \to C)\,,
    \]
  where $\mu$ is $\langle\ul{\pr_X},\ul{\pr_Y}\rangle$.

@@

  By the definition of $\mu$, we know that if the $i$-th pair of moves in $\t\vert_{\ul X,\ul Y}$ is $qx$ for $x\in X$, then the $i$-th pair of moves in $t\vert_{\ul{X\times Y}}$ is $q(x,y_0)$ for some $y_0$, and if the $i$-th pair of moves in $\t\vert_{\ul X,\ul Y}$ is $qy$ for $y\in Y$, then the $i$-th pair of mvoes in $t\vert_{\ul{X\times Y}}$ is $q(x_0,y)$ for some $x_0\in X$.
  Moreover, if $t'$ is some other sequence such that $t'/s$ and such that $t'$ differs only from $t$ in the choice of the `irrelevant' moves $x_0,y_0$, then there is some $S'\in\sigma\|(\ul X \to \tau)\|(\Lambda\inv;\mu)$ such that $S\vert_{A,\ul{X\times Y}\to C}=t'$ and $S\vert_{A,\ul X \to (\ul Y \to C)}=\t$.
  Then the combined probability of all such sequences is given by
  \[
    \left[ \prod_{i=1}^{|u|-1} \bP(V = u^{(i)}) \prod_{i=0}^{|v|-1} \bP(W = v^{(i)}) \;\middle|\; \mbox{\pbox{\textwidth}{where\\$\t\vert_{\ul X} = qu$ \\ $\t\vert_{\ul Y} = qv$}} \right]\,,
    \]
  and we can combine this with our calculations above to get the desired result.
\end{proof}

@@

This is the approach taken by Danos and Harmer in \cite{DanosHarmer} -- the original game semantics for a probabilistic variant of Algol.  
Danos and Harmer define an arena as we do, but define a probabilistic strategy on an arena $A$ to be given by a function
\[
  \sigma \from \L_A^{\text{even}} \to [0,1]
  \]
such that for any even-length $s$, and any extension $sa$, we have
\[
  \sigma(s) \ge \sum_{sab\in\L_A} \sigma(sab)\,.
  \]for any even-length $s$, and any extension $sa$, we have
\[
  \sigma(s) \ge \sum_{sab\in\L_A} \sigma(sab)\,.
  \]
These quantites $\sigma(s)$ take the role of our $\bP(s\in\sigma)$, but now they are part of the \emph{definition} of the strategy $\sigma$, rather than being a calculated quantity.

>>>>

@@

\newcommand{\IAXX}{{IA${}_{\X}$}\xspace}
\section{The language \IAXX}

@@

Let $\G$ be a model of Idealized Algol as described above, and suppose that there is an oplax monoidal functor $\Set \to \G$ that is used to interpret datatypes.  
We will use an underline to indicate this functor; so, for example, the object of $\G$ that is used to denote the natural number type is written $\ul\bN$.

By our description of $\G/\oppcat \X$ as a lax colimit in $\Cat$ (i.e., Corollary \ref{CorTheConstructionUniversalProperty}), we have a natural functor $J\from \G\to\G/\oppcat \X$ and a natural transformation $\phi_{p,a}\from J(jp\to a) \to Ja$.  
Our denotational semantics of \IAXX is then given in the category $\G/\oppcat X$ as follows.  
The denotation of any type $T$ of Idealized Algol is given by $J(\deno{T}_\G)$, where $\deno{T}_\G$ is the original denotation in $\G$.
The denotation of any sequent $\Gamma\ts M$ is given by $\deno{\Gamma\ts M} = J(\deno{\Gamma\ts M}_\G)$, where $\deno{-}_\G$ is the original denotation in $\G$.
The denotation of $\choose_p$ is given by the morphism $\omega_p \from 1 \to j(p)$ given by the composite
\[
  I \xrightarrow{\Lambda(\id_{Jjp})} (Jjp \to Jjp)  \to J(jp \to jp) \xrightarrow{\phi_{p,jp}} Jjp\,.
  \]

@@

Recall that, in the definition of the category $\G/\oppcat\X$, the \Mellies morphisms are left unchanged by precomposing with a morphism in the image of the functor $j$; therefore, if $M\from T$ is a closed term of \IAXX referring to $\choose_{p_1}$, $\cdots$, $\choose_{p_n}$, then we may write the denotation of $M$ as the composite
\[
  j(N) \xrightarrow{jf} j(p_1 \tensor \cdots \tensor p_n) \xrightarrow{m^j} j(p_1) \times \cdots \times j(p_n) \xrightarrow{\deno{x_1,\cdots,x_n\ts M[x_i/\choose_{p_i}]}_\G} \deno T\,;
  \]
i.e., as
\[
  j(N) \xrightarrow{\langle \pi_1,\cdots,\pi_n\rangle} j(p_1) \times \cdots \times j(p_n) \xrightarrow{\deno{x_1,\cdots,x_n\ts M[x_i/\choose_{p_i}]}_\G} \deno T\,.
  \]

@@

\begin{lemma}
  Let
  \[
    \inferrule{\Gamma,s^{(0)}\ts M_1\converges_{u_1} c_1,s^{(1)} \\ \cdots \\ \Gamma,s^{(n-1)}\ts M_n\converges_{u_n} c_n,s^{(n)}}
    {\Gamma,s^{(0)} \ts M \converges_{u_1\cat \cdots \cat u_n} c,s^{(n)}}
    \]
  be an inference of IA${}_{j(N)}$, where every instance of $\ask_{j(N)}$ occurs as part of some term $\neww (\lambda v.v\gets \ask_{j(N)};\Pi_i \oc v)$, and suppose that $M\ne \neww(\lambda v.v\gets\ask_{j(N)};\Pi_j\oc v)$ for any $j$.  
  Suppose we have sequences $U_1,\cdots,U_n$ such that $u_k$ covers $U_k$ for $k=1,\cdots,n$.
  Then
  \[
    \inferrule{\Gamma,s^{(0)}\ts M_1[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_1} c_1,s^{(1)} \\ \cdots \\ \Gamma,s^{(n-1)}\ts M_n[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_n} c_n,s^{(n)}}
    {\Gamma,s^{(0)} \ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_{U_1\cat \cdots \cat U_n} c,s^{(n)}}
    \]
  is a valid inference of \IAXX.
  \label{LemFourthSubstitution}
\end{lemma}
\begin{proof}
  As in Lemma \ref{LemSecondSubstitution}, we can prove this by looking at cases.  
  For example, consider the sequencing rule
  \[
    \inferrule{\Gamma,s\ts M\converges_u \skipp,s' \\ \Gamma,s' \ts N \converges_v c,s''}{\Gamma,s\ts M;N \converges c,s''}\,.
    \]
  We have
  \begin{IEEEeqnarray*}{Cl}
    &(M;N)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \\
    = &M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)];\\&N[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)]\,,
  \end{IEEEeqnarray*}
  and so we certainly get a rule
  \[
    \inferrule*{\Gamma,s\ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)]\converges_U \skipp,s' \\ \Gamma,s' \ts N[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_V c,s''}{\Gamma,s\ts (M;N)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges c,s''}\,.
    \]
  The only case where we need to be careful is for the $\neww$ rule:
  \[
    \inferrule*{\Gamma,x,(s\vert x\mapsto 0)\ts M\converges_u c,(s'\vert x\mapsto n)}
    {\Gamma,s\ts \neww \lambda x.M \converges_u c,s'}\,.
    \]
  If $\neww\lambda x.M \ne \neww(\lambda v.v\gets \ask_{j(N)};\Pi_j \oc v)$, then we have
  \begin{IEEEeqnarray*}{Cl}
    &(\neww \lambda x.M)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \\
    = & \neww \lambda x. (M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)])\,.
  \end{IEEEeqnarray*}
  Then we can apply the rule
  \[
    \inferrule*{\Gamma,x,(s\vert x\mapsto 0)\ts M[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges_U c,(s'\vert x\mapsto n)}
    {\Gamma,s\ts (\neww\lambda x.M)[\choose_{p_i}/\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i\oc v)] \converges U c,s'}.\qedhere
    \]
\end{proof}

@@

\begin{lemma}
  Let $\Gamma,y_1\from j(p_1),\cdots,y_n\from j(p_n) \ts M \from T$ be a term-in-context of ordinary Idealized Algol, where $\Gamma$ is a $\Var$-context.  
  Let $U$ be a sequence and let $N,\pi_i,\Pi_i$ be as above.
  Suppose that there exists some sequence $u\in j(N)^*$ such that $u$ covers $U$ and such that
  \[
    \Gamma,s\ts M[\neww(\lambda v.v\gets \ask_{j(N)};\Pi_i \oc v)/y_i]\converges_u c,s'\,.
    \]
  Then
  \[
    \Gamma,s\ts M[\choose_{p_i}/y_i] \converges_U c,s'\,.
    \]
  \label{LemAdequacyIaxx}
\end{lemma}
\begin{proof}
  Induction on the derivation.
  Suppose that $M$ is not one of the $y_i$; then $M[\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)/y_i]$ is not equal to $\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)$.
  Moreover, every instance of $\ask_{j(N)}$ in $M$ occurs as part of an expression of the form $\neww(\lambda v.v\gets\ask_{j(N)};\Pi_i \oc v)$, and so we win by Lemma \ref{LemFourthSubstitution} and the inductive hypothesis.

@@

Given a closed term $M\from\com$ of IA${}_\bB$, we define
\[
  \bP(M\converges) = \sum_{u\in\bB^*\colon M\converges_u\skipp} 2^{-|u|}\,,
  \]
since $2^{-|u|}$ is the probability of a particular sequence $u$ of $\true$ and $\false$ values occurring.
Here, the infinite sum means the supremum over all sums of finite subsets.
Similarly, given a Kleisli morphism $\sigma\from 1 \to \bC$ -- i.e., a morphism $\sigma\from 1 \to (\bB \to \bC)$ in $\G$, we can define
\[
  \bP(\sigma\downarrow) = \sum_{u\in\Acc(\sigma)} 2^{-|u|}\,.
  \]
Our Computational Adequacy result for \IAX (Propositions \ref{PropKleisliSoundness} and \ref{PropKleisliAdequacy}) then gives us a Computational adequacy result for this model.

@@

This model is much easier to realize.  
What do we gain, then, from our more complicated original model?

The most obvious answer is that our original model allowed us to work with arbitrary probabilities, rather than using a fixed coin with probability $\frac12$.  
This is not such a great advantage as it might seem, since if $p\in[0,1]$ is any real number whose binary expansion is computable as a function $\bN \to \bB$, then we can simulate $\choose_p$ within the probabilistic version of IA${}_\bB$\footnote{Idea: build up a sequence of intervals $I_n$ of width $2^n$ by repeatedly flipping the coin and using the result to choose either the lower or the upper half of $I_{n-1}$.  
If at any point the upper bound of $I_n$ is less than or equal to $p$ (which can be checked by computing the first $n$ terms of the binary expansion of $p$), then return $\true$.  
If at any point the lower bound of $I_n$ is greater than or equal to $p$, return $\false$.}.

@@

\begin{proposition}
  Let $\sigma\from A \to B$, $\tau\from B \to C$ be morphisms in the category $\G/\oppcat{(\Rv_{\Omega}^{FS})}$.  
  Let $s$ be a legal play in $A \to C$.
  Then
  \[
    \bP(s\in\sigma;\tau) = \sum_{\s\in\wit_B(s)}\bP(\s\vert_{A,B}\in\sigma)\bP(\s\vert_{B,C}\in\tau)\,.
    \]
  \label{PropCompositionProb}
\end{proposition}
\begin{proof}
  Suppose that $\sigma$ and $\tau$ are given by (equivalence classes of) pairs
  \begin{mathpar}
    (V\from\Omega\to X, \tilde \sigma \from A \to (\ul X \to B))
    \and
    (W\from\Omega\to Y, \tilde \tau \from B \to (\ul Y \to C))\,,
  \end{mathpar}
  where $V$ and $W$ are random variables and $\tilde \sigma,\tilde\tau$ are strategies in $\G$.
  Then the composition $\sigma;\tau$ in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$ is given by $V\tensor W\from\Omega \to X\times Y$, together with the \Mellies composition
  \begin{mathpar}
    A \xrightarrow{\tilde\sigma}
    (\ul{X} \to B) \xrightarrow{\ul{X} \to \tilde\tau}
    (\ul X \to (\ul Y \to C)) \to
    ((\ul X \times \ul Y) \to C) \xrightarrow{\mu\to C}
    (\ul{X\times Y} \to C)\,,
  \end{mathpar}
  where $\mu$ is $\langle\ul{\pr_X},\ul{\pr_Y}\rangle$.

@@

  By the definition of $\mu$, we know that if the $i$-th pair of moves in $\t\vert_{\ul X,\ul Y}$ is $qx$ for $x\in X$, then the $i$-th pair of moves in $t\vert_{\ul{X\times Y}}$ is $q(x,y_0)$ for some $y_0$, and if the $i$-th pair of moves in $\t\vert_{\ul X,\ul Y}$ is $qy$ for $y\in Y$, then the $i$-th pair of moves in $t\vert_{\ul{X\times Y}}$ is $q(x_0,y)$ for some $x_0\in X$.
  Moreover, if $t'$ is some other sequence such that $t'/s$ and such that $t'$ differs only from $t$ in the choice of the `irrelevant' moves $x_0,y_0$, then there is some $S'\in\sigma\|(\ul X \to \tau)\|(\Lambda\inv;\mu)$ such that $S\vert_{A,\ul{X\times Y}\to C}=t'$ and $S\vert_{A,\ul X \to (\ul Y \to C)}=\t$.
  Then the combined probability of all such sequences is given by
  \[
    \left[ \prod_{i=1}^{|u|-1} \bP(V = u^{(i)}) \prod_{i=0}^{|v|-1} \bP(W = v^{(i)}) \;\middle|\; \mbox{\pbox{\textwidth}{where\\$\t\vert_{\ul X} = qu$ \\ $\t\vert_{\ul Y} = qv$}} \right]\,,
    \]
  and we can combine this with our calculations above to get the desired result.
\end{proof}

@@

This is the approach taken by Danos and Harmer in \cite{DanosHarmer} -- the original game semantics for a probabilistic variant of Algol.  
Danos and Harmer define an arena as we do, but define a probabilistic strategy on an arena $A$ to be given by a function
\[
  \sigma \from \L_A^{\text{even}} \to [0,1]
  \]
such that for any even-length $s$, and any extension $sa$, we have
\[
  \sigma(s) \ge \sum_{sab\in\L_A} \sigma(sab)\,.
  \]for any even-length $s$, and any extension $sa$, we have
\[
  \sigma(s) \ge \sum_{sab\in\L_A} \sigma(sab)\,.
  \]
These quantities $\sigma(s)$ take the role of our $\bP(s\in\sigma)$, but now they are part of the \emph{definition} of the strategy $\sigma$, rather than being a calculated quantity.
